---
title: "Northville-Placid Trail"
author: "Nathan Hoteling"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(lubridate)

library(ggridges)
library(rgdal)
library(raster)
library(magick)    # image manipulations
library(plotKML)   # Read the KML?
library(rayshader) # elevation matrix stuff
library(RcppRoll)  # smoothing ele profile for ridgeline

library(r2d3)
library(jsonlite)
```

## Data

There are basically three different data files used here:  
1. The elevation profile, obtained from [Earth Explorer](https://earthexplorer.usgs.gov)  
2. Northville-Placid Trail coordinates, obtained from [The Hiking Project](https://www.hikingproject.com/trail/7004949/npt-northville-placid-trail)  
3. Adirondack Park Boundary, obtained from [Adirondack Park Agency](https://apa.ny.gov/gis/)  

The Earth Explorer data were downloaded as separate files and merged together, with `raster` and `rgdal` packages in R.  The NPT coordinates were read directly from the downloaded GPX file, and the ADK park boundary was read directly from the KML.  An image of the NPT trail marker was downloaded from somewhere to spice up the map display.

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Get the code for the appropriate UTM zone
# See README file for how we get the epsg code
epsg <- make_EPSG()
prj4 <- epsg$prj4[ epsg$code==2032 ]
crd.new <- CRS(prj4[1])


#-----------------------------------
#  Elevation profile data comes from Geotif files
#  downloaded from  https://earthexplorer.usgs.gov
#  and merged together with gdal command-line tools.
#  (see README file for more details)
#

tifname <- "data/adk_3arc_merged.tif"    # 17.3 MB  3601x2401 pixels  ***
localtif = raster::raster(tifname)
elmat <- raster_to_matrix(localtif)

# NPT logo image
imgNPT <- as.raster(image_read("data/npt-image-v02.png"))

#-------------------------------------
#  Adirondack Park Border
#
shp.adk <- readOGR("data/Adirondack_Park_Boundary.kml")

#-----------------------------------
#  Northville-Placid Trail coordinates
#  GPX file downloaded from hikingproject(?)
#
route <- plotKML::readGPX("data/npt-northville-placid-trail.gpx")
npt <- data.frame("lon"=route$tracks[[1]]$'NPT (Northville-Placid Trail)'$lon,
                  "lat"=route$tracks[[1]]$'NPT (Northville-Placid Trail)'$lat,
                  "ele"=as.numeric(route$tracks[[1]]$'NPT (Northville-Placid Trail)'$ele),
                  "time"=as.character(floor_date(ymd_hms(route$tracks[[1]]$'NPT (Northville-Placid Trail)'$time), unit="minutes")))
```

## Static Map with R


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#------------------------------------------
# Create ridgeplot from Geotif
# --> Part of this comes from rayshader tutorial
# by elizabeth easter
#
xmin <- localtif@extent@xmin
ymin <- localtif@extent@ymin
xmax <- localtif@extent@xmax
ymax <- localtif@extent@ymax
xh <- (xmax-xmin)/2
yh <- (ymax-ymin)/2

df.pts <- SpatialPoints(cbind(c(xmin,xmax),c(ymin,ymax)), proj4string=CRS(proj4string(localtif)))
df.pts2 <- spTransform(df.pts,crd.new)
crd <- coordinates(df.pts2)
asp.rto <- (crd[4]-crd[3]) / (crd[2]-crd[1])

# Use roll_mean() from RcppRoll package
N <- 10  # number of points to include for rolling average
r_elmat <- roll_mean(elmat, n=N,fill=NA,align="center")
t_elmat <- roll_mean(t(r_elmat),n=N,fill=NA,align="center")
x_telmat <- t_elmat[seq(from=1,to=nrow(t_elmat),by=N),seq(from=1,to=ncol(t_elmat),by=N)]

lat_val <- rev((1:localtif@nrows)*res(localtif)[1]+ymin)
r_lat_val <- roll_mean(lat_val,n=N,fill=NA,align="center")
x_lat_val <- r_lat_val[seq(from=1,to=length(r_lat_val),by=N)]
lon_val <- (1:localtif@ncols)*res(localtif)[2]+xmin
r_lon_val <- roll_mean(lon_val,n=N,fill=NA,align="center")
x_lon_val <- r_lon_val[seq(from=1,to=length(r_lon_val),by=N)]

df.telmat <- data.frame(x_telmat)
names(df.telmat) <- x_lon_val
df.telmat$lat <- x_lat_val
md.telmat <- melt(df.telmat, id=c("lat"))

df.adk <- md.telmat[complete.cases(md.telmat),]
df.adk$lon <- as.numeric(as.character(df.adk$variable))
df.adk$value2 <- df.adk$value
df.adk$value2[df.adk$value==0] <- NA
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}
#--------------------------------------------
# Make plot
#



rstr.xmin <- -75.92
rstr.xmax <- -75.62
rstr.ymin <- 44.72
rstr.ymax <- rstr.ymin+(rstr.xmax-rstr.xmin)*0.75 #asp.rto

banner <- data.frame(id="banner",x=c(-75.92,-75.92,-74.72,-74.72),
y=c(44.72,rstr.ymax+0.02,rstr.ymax+0.02,44.72))

p.rdg <- ggplot(df.adk) +
  geom_density_ridges(aes(x=lon,y=lat,height=value2,group=lat),
                      stat="identity", scale=8, color="grey45",size=0.3,fill=NA) +
  geom_polygon(data=banner, aes(x=x,y=y),fill="#fff1e5",color="grey65",size=1.0) +
  geom_path(data=shp.adk, aes(x=long,y=lat,group=group),
            color="darkgreen",size=1.0,alpha=0.3) +
  geom_path(data=npt, aes(x=lon,y=lat), color="steelblue",size=0.65,alpha=0.8) +
  annotation_raster(imgNPT,rstr.xmin,rstr.xmax,rstr.ymin,rstr.ymax,interpolate=TRUE) +
  annotate("text",x=rstr.xmin+0.3,y=rstr.ymin,label="Northville-Placid\nTrail",
           color="grey20",hjust=0.0,vjust=-0.18,size=4.5,family="Optima",fontface="bold") +
  theme_void() +
  theme(panel.background=element_rect(fill="#fff1e5"),
        plot.background=element_rect(fill="#fff1e5"),
        aspect.ratio=asp.rto)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, out.width="90%", eval=TRUE}
p.rdg
```
<br>
<br>

## Interactive Map with D3

```{r, echo=FALSE}
d.data <- list(adk=df.adk, npt=npt)
dd <- jsonlite::toJSON(d.data)
write(dd,"data/npt-map.json")
r2d3(data=dd, 
     script="js/npt-map.js", 
     container="div",
     dependencies = c("js/d3-simple-slider.min.js"),
     height=400)
```

<br>
<br>